Recalibrating Gravitational Wave Phenomenological Waveform Model

### 2023-06-23 src/tex/ms.tex
Finished section 3

With the purpose of improving downstream tasks such as parameter estimation in
mind, the more relevant metric of improvement is the distribution of improvement
in mismatch over the entire validation dataset. Figure~\ref{fig:q148} shows the
distribution of log-mismatches for the validation waveforms before and after the
optimization procedure. Here we show results using a constant PSD in our loss
function. One can see the distribution of the optimized waveform is skewed toward lower mismatches,
with the peak of the distribution being shifted by approximately an order of
magnitude, and a median mismatch being reduced by 50\% (see vertical dotted lines). When using
$\mathcal{L}_{\mathrm{nl}}$, we observe a less pronounced improvement with a 22.9\%
decrease in the median.

@@

Note that the IMRPhenomD model was initally tested using the
{\zdethp} weighted mismatch. We would therefore like to examine whether using the {\zdethp} PSD in our loss function
could lead to an improved to mismatch. Peforming the optimization again, we find no significant difference
between the results using the two PSDs in the distribution of mismatches.

@@

To investigate the performance of recalibration over the source parameter space,
we plot the improvement of the log-mismatch as a function of the parameter space
$q-\chi_{\mathrm{PN}}$ in Fig. \ref{fig:ps_q148_qchi}. Red points indicate that the waveform
is improved by the optimization procedure, while blue points indicate that the waveform mismatch
\textit{increases} during optimization.
We can see that waveforms with $q\le4$ show the most consistent average improvement.
This is likely due to the better coverage of training waveforms in that part of the parameter space (see Fig.~\ref{fig:intrin_space}).
On the spin axis, we can see the waveforms with
$\chi_{\textrm{PN}}\sim0$ show the most consistent improvement. 
When we move away from $\chi_{\textrm{PN}}\sim0$, the improvement fluctuates but exhibits an overall trend. 
This is particularly true in the $q\le4$ region, where we see a consistent
improvement of the waveform for $\chi_{\textrm{PN}}<0$. 
Since the original waveform was developed for aligned-spin systems, the waveforms
with opposite sign $\chi_{1,2}$ are likely less well-fit \kw{Verify this.}.
To show this, we plot the parameter space $\chi_1 - \chi_2$ in Fig.~\ref{fig:ps_q148}. 
Waveforms along the diagonal axis, $\chi_1\sim\chi_2$, show good mismatch improvements as discussed
above. Meanwhile, the top-left ($\chi_1<0<\chi_2$) and bottom-right
($\chi_1>0>\chi_2$) regions respond to the optimization differently. In the top-left
region, waveforms generally improve slightly after optimization. However,
waveforms in the bottom-right region do not improve after optimization.

@@

We divided the parameter space into four regions to analyze the effect of the
recalibration procedure on each region separately
(Fig.~\ref{fig:ps_q148_quadrant}). The training waveforms used for fitting in
this scenario are listed in Tab.~\ref{tab:quadrants} and the loss functions 
are calculated using simple average of the mismatches, $\mathcal{L}_{\mathrm{ave}}$. The top-left and
bottom-right regions have limited data for $q>4$, hence the results are only
valid up to $q\lesssim4$. From Fig.~\ref{fig:ps_q148_quadrant}, we observe that
equal-spin waveforms lying on the diagonal improve the most. The
improvement above the diagonal is also significantly more than when optimizing using
all waveforms at once. This can be seen clearly in the top panel of Fig.~\ref{fig:all_quadrants}, where we see
that the optimized histogram shifts to lower mismatches.\footnote{
	Note that the top panel of Fig.~\ref{fig:all_quadrants} specifically shows the
	distribution of mismatches for the top-left quadrant of Fig.~\ref{fig:ps_q148_quadrant}.
} On the other hand, the
improvement below the diagonal is negligible, indicating that our method struggles to
further improve on top of the original IMRPhenomD waveform. 
Interestingly, we can see in the bottom-right quadrant that the waveform does not show a significant change
to the overall mismatch (see bottom panel of Fig.~\ref{fig:all_quadrants}). 
This suggests that the IMRPhenomD waveform ansatz fits waveforms with
$\chi_1>0$ and $\chi_2<0$ less well.

>>>>

With the purpose of improving downstream tasks such as parameter estimation in
mind, the more relevant metric of improvement is the distribution of improvement
in mismatch over the entire validation dataset. Figure~\ref{fig:q148} shows the
distribution of log-mismatches for the validation waveforms before and after the
optimization procedure. Here we show results using a constant PSD in our loss
function. One can see the distribution of the optimized waveform is skewed toward lower mismatches,
with the peak of the distribution being shifted by approximately an order of
magnitude, and the median mismatch being reduced by 50\% (see vertical dotted lines). When using
$\mathcal{L}_{\mathrm{nl}}$, we observe a less pronounced improvement with a 22.9\%
decrease in the median.

@@

Note that the performace of the IMRPhenomD model was initally tested using the
{\zdethp} weighted mismatch.\footnote{
	Note, however, that the mismatch was never directly used during the calibration process~\citep{khan2016frequency, khan2019phenomenological}.
} We would therefore like to examine whether using the {\zdethp} PSD in our loss function
could lead to an improved to mismatch. Peforming the optimization again, we find no significant difference
between the results using the two PSDs in the distribution of mismatches.

@@

To investigate the performance of recalibration over the source parameter space,
we plot the improvement of the log-mismatch as a function of the parameter space
$q-\chi_{\mathrm{PN}}$ in Fig. \ref{fig:ps_q148_qchi}. Red points indicate that the waveform
is improved by the optimization procedure, while blue points indicate that the waveform mismatch
\textit{increases} during optimization.
We can see that waveforms with $q\le4$ show the most consistent average improvement.
This is likely due to the better coverage of training waveforms in that part of the parameter space (see Fig.~\ref{fig:intrin_space}).
On the spin axis, we can see that the waveforms with
$\chi_{\textrm{PN}}\sim0$ show the most consistent improvement. 
When we move away from $\chi_{\textrm{PN}}\sim0$, the improvement fluctuates but exhibits an overall trend. 
This is particularly true in the $q\le4$ region, where we see a consistent
improvement of the waveform for $\chi_{\textrm{PN}}<0$. 
Since the original model was developed for aligned-spin systems, NR waveforms
with opposite sign $\chi_{1,2}$ are likely less well-fit \kw{Verify this.}.
To show this, we plot the parameter space $\chi_1 - \chi_2$ in Fig.~\ref{fig:ps_q148}. 
Points along the diagonal axis, $\chi_1\sim\chi_2$, show good mismatch improvements as discussed
above. Meanwhile, the top-left ($\chi_1<0<\chi_2$) and bottom-right
($\chi_1>0>\chi_2$) regions respond to the optimization differently. In the top-left
region, the waveform generally improves after optimization. However,
in the bottom-right region, the waveform does not improve after optimization.

@@

We divided the parameter space into four regions to analyze the effect of the
recalibration procedure on each region separately
(Fig.~\ref{fig:ps_q148_quadrant}). The training waveforms used for fitting in
this scenario are listed in Tab.~\ref{tab:quadrants} and the loss functions 
are calculated using a simple average of the mismatches, $\mathcal{L}_{\mathrm{ave}}$. The top-left and
bottom-right regions have limited data for $q>4$, hence the results are only
valid up to $q\lesssim4$. From Fig.~\ref{fig:ps_q148_quadrant}, we observe that
equal-spin waveforms lying on the diagonal improve the most. The
improvement above the diagonal is also significantly more than when optimizing using
all waveforms at once. This can be seen clearly in the top panel of Fig.~\ref{fig:all_quadrants}, where we see
that the optimized histogram shifts to lower mismatches.\footnote{
	Note that the top panel of Fig.~\ref{fig:all_quadrants} specifically shows the
	distribution of mismatches for the top-left quadrant of Fig.~\ref{fig:ps_q148_quadrant}.
} On the other hand, the
improvement below the diagonal is negligible, indicating that our method struggles to
further improve over the original IMRPhenomD waveform. 
Interestingly, we can see in the bottom-right quadrant that the waveform does not show a significant change
to the overall mismatch (see bottom panel of Fig.~\ref{fig:all_quadrants}). 
This suggests that the IMRPhenomD waveform ansatz fits waveforms with
$\chi_1>0$ and $\chi_2<0$ less well.
